---
title: "Expedia Hotel Recommendation - Initial Analysis"
author: "Anandan Balaji"
date: "12 May 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This R Markdown doucment illustrates the different plots for some of the data fields related to the Capstone Project - Expedia Hotel Recommendation.

The train data is huge - roughly 4GB. So, inorder to load and render fast, first 25K records were extracted and the following plots are generated for initial data analysis process.

```{r train dataset}
# read 25 K
train_df <- read.csv(file = "train.25000.csv")

library(ggplot2)
```

## site_name

```{r plot0, echo=FALSE}
ggplot(train_df, aes(site_name)) + geom_bar(fill = "purple")
```

## posa_continent - current user location
```{r plot1, echo=FALSE}
ggplot(train_df, aes(posa_continent)) + geom_bar(fill = "brown")
```

## hotel_continent - destination location

```{r plot2, echo=FALSE}
ggplot(train_df, aes(hotel_continent)) + geom_bar(fill = "green")
```

## user_id

```{r plot3, echo=FALSE}
ggplot(train_df, aes(user_id)) + geom_bar(fill = "blue")
```

## is_mobile 

```{r plot4, echo=FALSE}
ggplot(train_df, aes(is_mobile)) + geom_bar(fill = "orange")
```

## is_package - hotel / hotel + air
 
```{r plot5, echo=FALSE}
ggplot(train_df, aes(is_package)) + geom_bar(fill = "green")
```

## srch_adults_cnt

```{r plot6, echo=FALSE}
ggplot(train_df, aes(srch_adults_cnt)) + geom_bar(fill = "red")
```

## srch_children_cnt
 
```{r plot7, echo=FALSE}
ggplot(train_df, aes(srch_children_cnt)) + geom_bar(fill = "red")
```

## srch_rm_cnt
 
```{r plot8, echo=FALSE}
ggplot(train_df, aes(srch_rm_cnt)) + geom_bar(fill = "red")
```

## hotel_cluster

```{r plot9, echo=FALSE}
ggplot(train_df, aes(hotel_cluster)) + geom_bar(fill = "blue") 
```
 