---
title: "expediaEDA_ext"
author: "Anandan Balaji"
date: "18 May 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Expedia Hotel Recommendation - Data Anyalysis

This document shows the charts related to different fields in the training data set.


```{r }
# read 25 K
#system.time(train_df <- read.csv(file = "train.25000.csv"))
library(data.table)
#system.time(train_df <- fread("train.25000.csv"))
system.time(train_df <- fread("train.csv"))
library(ggplot2)
```

 
## posa_continent (user location) and hotel_continent (desitnation location)

The below chart shows that the people from continent #3 travel a lot and prefer visiting continent #2.

```{r plot1 }
ggplot(train_df, aes(posa_continent, fill = factor(hotel_continent))) + geom_bar(position = "dodge")
```

 ## srch_adults_cnt, srch_children_cnt
```{r}
keys <- c("date_time", "user_id")
setkeyv(train_df, keys)

#adult_df <- train_df[keys]
adult_df <- train_df[keys]
adult_df$count <- train_df$srch_adults_cnt
adult_df$type <- "Adult"

child_df <- train_df[keys]
child_df$count <- train_df$srch_children_cnt
child_df$type <- "Children"

adult_child_df <- rbind(adult_df, child_df)

#ggplot(adult_child_df, aes(x = user_id, y = count, col = factor(type))) +
#  geom_line()

# how to change the scale in x-axis to 0, 1, 2, 3 etc.?
ggplot(adult_child_df, aes(count, col = factor(type))) + geom_freqpoly()
```

## source and destination country
```{r}

src_df <- train_df[keys]
src_df$country <- train_df$user_location_country
src_df$ctype <- "User Country"

dest_df <- train_df[keys]
dest_df$country <- train_df$hotel_country
dest_df$ctype <- "Dest Country"

country_df <- rbind(src_df, dest_df)

ggplot(country_df, aes(country, col = factor(ctype))) + 
  geom_freqpoly()
```

## number of days stay : srch_co - srch_ci (checkout - checkin date)

```{r}
ggplot(train_df, 
       aes(factor(abs(as.Date(as.character(train_df$srch_co), format = "%Y-%m-%d") -                           as.Date(as.character(train_df$srch_ci), format = "%Y-%m-%d"))))) +
        geom_bar(col = "blue") +
        xlab("hotel stay in days")
```


## hotel_continent & hotel_cluster

There are more number of hotel_clusters in the continent #2 and #6.

```{r}
ggplot(train_df, 
       aes(x = hotel_cluster, y = hotel_continent, col = factor(hotel_continent))) +       # to avoid over plotting    
  geom_jitter(alpha = 0.7) +
      geom_smooth(method = "lm", se = F)
```
